\section{Shortest Paths}\label{sec:1}

\subsection{Preliminaries on Graphs}\label{subsec:1.1}
An {\bf (undirected) graph} $G$ is a pair $(V, E)$, where $E$ is a set of 
unordered pairs of elements in $V$. The elements of $V$ are called {\bf vertices}
or {\bf nodes}; the elements of $E$ are called {\bf edges}. 

Let $u, v \in V$ and let $e = uv \in E$ be an edge. 
\begin{itemize}
    \item We say that $e$ is {\bf incident} to $u$ and $v$. 
    \item The vertices $u$ and $v$ are said to be {\bf adjacent}.
    \item We call $u$ and $v$ the {\bf endpoints} of $e$. 
\end{itemize}
By default, we assume that there are no parallel edges (i.e. two
edges $e = uv$ and $e' = u'v'$ in $E$ with $\{u, v\} = \{u', v'\}$) 
and no loops (i.e. an edge $e = uv \in E$ with $u = v$).

For distinct $u, v \in V$, a {\bf $u, v$-path} is a sequence of vertices 
$w_1, \dots, w_k$ such that $w_1 = u$, $w_k = v$, and $w_i w_{i+1} \in E$ 
for all $i = 1, \dots, k-1$.  

For example, consider the following graph $G = (V, E)$ with 
vertices $V = \{v_1, v_2, v_3, v_4\}$ and edges $E = \{v_1v_2, v_1v_4, v_2v_3, 
v_2v_4, v_3v_4\}$. 

\begin{center}
    \begin{tikzpicture}[node distance={15mm}, main/.style = {draw, circle}] 
        \node[main] (1) {$v_1$}; 
        \node[main] (2) [right of=1] {$v_2$};
        \node[main] (3) [below of=2] {$v_3$}; 
        \node[main] (4) [left of=3] {$v_4$};

        \draw (1) -- (2);
        \draw [draw=red] (1) -- (4);
        \draw [draw=red] (2) -- (3);
        \draw (2) to [out=315, in=315, looseness=2] (4);
        \draw [draw=red] (3) -- (4);
    \end{tikzpicture} 
\end{center}
\vspace{-0.5cm}

The lines in red form a $v_1, v_2$-path, namely $v_1, v_4, v_3, v_2$. 
Another $v_1, v_2$-path can be obtained by simply traversing the edge $v_1v_2$. 

A {\bf cycle} in $G$ is a sequence of vertices $w_1, \dots, w_{k+1}$ 
such that $w_i w_{i+1} \in E$ for all $i = 1, \dots, k$, the vertices 
$w_1, \dots, w_k$ are all distinct, and $w_1 = w_{k+1}$.

Finally, a graph $G$ is {\bf connected} if for any pair of distinct vertices 
$u, v \in V$, there exists a $u, v$-path in $G$. 

\subsection{Shortest Paths Problem}\label{subsec:1.2}
Given a \emph{directed} graph $G = (V, E)$ with edge lengths $\ell_e \geq 0$
for each $e \in E$ and a distinguished start vertex $s \in V$, we wish 
to find shortest paths from $s$ to every other vertex in $V$. Note that 
when we work with directed graphs, we will denote the directed edges 
with $(v_1, v_2)$ as opposed to $v_1 v_2$ in the case of undirected graphs, where 
the order of the vertices did not matter. 

The {\bf length} of a path $P$ given by the sequence $w_1, \dots, w_k$ 
is given by 
\[ \ell(P) := \sum_{i=1}^{k-1} \ell_{(w_i, w_{i+1})} = \sum_{e\in P} \ell_e, \] 
where the second sum makes sense because there are no parallel edges. 
Then the {\bf shortest-path distance} from $s$ to a vertex $u \in V$ is 
defined to be
\[ d(u) := \min_{\text{$s,u$-paths $P$}} \ell(P). \] 
For example, we can consider the following instance of an undirected graph 
with given edge lengths and starting vertex $s = v_1$. 

\begin{center}
    \begin{tikzpicture}[node distance={30mm}, main/.style = {draw, circle}] 
        \node[main] (1) {$v_1$}; 
        \node[main] (2) [below left of=1] {$v_2$};
        \node[main] (3) [below right of=1] {$v_3$}; 
        \node[main] (4) [below left of=2] {$v_4$};
        \node[main] (5) [below left of=3] {$v_5$};
        \node[main] (6) [below right of=3] {$v_6$};

        \draw[->] (1) -- node[midway, above left] {1} (2);
        \draw[->] (1) -- node[midway, above right] {2} (3);
        \draw[->] (1) -- node[midway, above right] {4} (5);
        \draw[->] (2) -- node[midway, above left] {4} (4);
        \draw[->] (2) -- node[midway, above right] {3} (5);
        \draw[->] (3) -- node[midway, above left] {1} (5);
        \draw[->] (3) -- node[midway, above right] {1} (6);
        \draw[->] (5) -- node[midway, above] {2} (4);
        \draw[->] (6) to [out=225, in=315, looseness=0.5] node[midway, above] {1} (4);
    \end{tikzpicture} 
\end{center}
\vspace{-0.5cm}
In this case, we have $d(v_2) = 1$, since the only possible path from 
$v_1$ to $v_2$ is by taking the edge $(v_1, v_2)$. There are multiple
paths from $v_1$ to $v_5$; the shortest one is $v_1, v_3, v_5$ giving 
$d(v_5) = 3$. 

Note that we always set $d(s) = 0$. We now make some observations: 
\begin{enumerate}[(i)]
    \item If $(u, v) \in E$, then $d(v) \leq d(u) + \ell_{(u,v)}$, since 
    such an $s, v$-path is always an option.
    \item For every $v \in V$ distinct from $s$, there exists $w \in V$ 
    such that $d(v) = d(w) + \ell_{(w, v)}$ and $(w, v) \in E$. This can 
    be seen by chopping off the last edge from a shortest path from $s$ to $v$.
\end{enumerate}

\subsection{Dijkstra's Algorithm}\label{subsec:1.3}
In 1959, Dijkstra came up with the following algorithm to solve the 
shortest paths problem. The main idea is to maintain a set $A \subseteq V$ 
of ``explored'' nodes; that is, a set of nodes for which we already know the 
shortest-path distances. We'll also maintain labels $d'(v)$ for $v \in 
V \setminus A$ with upper bounds on the shortest-path distances from $s$. 

\begin{mdframed}[
    linewidth=1pt,
    linecolor=black,
    bottomline=false,topline=false,rightline=false,
    innerrightmargin=0pt,innertopmargin=0pt,innerbottommargin=0pt,
    innerleftmargin=1em,% Distance between vertical rule & proof content
    skipabove=0.75\baselineskip
]
{\bf Input.} A directed graph $G = (V, E)$, edge lengths $\ell_e \geq 0$ for 
all $e \in E$, and a start vertex $v \in V$. 

{\bf Output.} For all $v \in V$, the length $d(v)$ for the shortest-path from 
$s$ to $v$.
\begin{enumerate}[leftmargin=1.75cm, label={Step \arabic*.}]
    \item {\bf Initialization.} Set $A \gets \{s\}$, $d(s) \gets 0$, and $d'(v) \gets \infty$ 
    for all $v \in V \setminus A$.

    \item While $A \neq V$:
    \begin{enumerate}[label={Step 2.\arabic*.}]
        \item {\bf Push down the upper bounds.} For each $v \in V \setminus A$, compute 
        \[ d'(v) \gets \min\left\{ d'(v), \min_{\substack{u\in A \\ (u, v) \in E}} 
        \{ d(u) + \ell_{(u, v)} \} \right\}. \] 
        \item {\bf Add a new vertex.} Set $w \gets \argmin_{v \in V \setminus A} d'(v)$, 
        $A \gets A \cup \{w\}$, and $d(w) \gets d'(w)$. 
    \end{enumerate}
\end{enumerate}
\end{mdframed}\vspace{-0.25cm}

Suppose that for each vertex $w \in V$, we keep track of the node $u$ 
determining its upper bound $d'(w)$. That is, the node $u$ is such that 
$(u, w) \in E$ and $d'(w) = d(u) + \ell_{(u, w)}$. Then at the end of the 
algorithm, a shortest path from $s$ to $w$ can be obtained as a shortest path 
from $s$ to $u$ adjoined with the edge $(u, w) \in E$. Moreover, these edges 
selected by Dijkstra's algorithm form an arborescence, which is a nice graph 
structure that we'll discuss more later. 

Next, let's prove the correctness of Dijkstra's algorithm. In particular, 
we need to show that for every $v \in V$, the distance from $s$ to $v$ 
is computed correctly. We'll assume that the graph is connected; that is, 
for every $v \in V$, there is an $s, v$-path in $G$. (Note that the 
algorithm won't terminate otherwise, but it can be adjusted to deal 
with this.)

\begin{pf}[correctness of Dijkstra's algorithm]
    We proceed by induction on $|A|$, and show that at each point in time, 
    $d(v)$ is computed correctly for all $v \in A$. The case where $|A| = 1$ 
    is clear because at the start of the algorithm, we initialize $A = \{s\}$ 
    with $d(s) = 0$, which is correct. 

    Assume that $d(v)$ is computed correctly for every $v \in A$ when 
    that $|A| = k$. Suppose that we are adding a new vertex $w$ to $A$ 
    in Step 2.2 of the algorithm. Consider the vertex $u \in A$ such that 
    $(u, w) \in E$ and 
    \[ d'(w) = d(u) + \ell_{(u, w)}. \] 
    Specifically, this is the vertex $u$ determining the upper bound 
    $d'(w)$ which we discussed in the paragraph following the description 
    of the algorithm. 

    For the sake of contradiction, assume that the distance from $s$ to $w$ 
    is not $d'(w)$. Let $P_u$ be a shortest path from $s$ to $u$, and 
    let $P'$ be a shortest path from $s$ to $w$. Then by our 
    assumption, we know that 
    \[ \ell(P') < \ell(P_u) + \ell_{(u, w)} = d'(w). \] 
    Now, let $x, y \in V$ be such that $(x, y) \in E$ lies on the shortest 
    path $P'$ from $s$ to $w$, with $x \in A$ and $y \in V \setminus A$. 
    (This exists because at some point, the path must exit $A$ to get 
    from $s$ to $w$.) Then we obtain 
    \[ d'(y) \leq d(x) + \ell_{(x,y)} \leq \ell(P') < \ell(P_u) 
    + \ell_{(u, w)} = d'(w), \] 
    where the first inequality is because of how $d'(y)$ is computed in 
    Step 2.1, and the second inequality is because the shortest path 
    from $x$ to $y$ adjoined with the edge $(x, y)$ is part of the path $P'$, 
    noting that $\ell_e \geq 0$ for all $e \in E$. But this contradicts 
    our choice of $w = \argmin_{v\in V \setminus A} \{d'(v)\}$ in Step 2.2 
    since $y \in V \setminus A$ but $d'(y) < d'(w)$. \qed
\end{pf}\vspace{-0.25cm}

The {\bf running time} of an algorithm is the number of elementary operations 
that the algorithm performs as a function of the input size. The {\bf input 
size} is the number of bits needed to specify the input. 
\begin{itemize}
    \item For example, in order to specify an integer $n \geq 0$ in binary, 
    we require about $\log_2(n)$ bits. 
    \item To specify a graph $G = (V, E)$ with integral edge lengths 
    $\ell_e \geq 0$ for each $e \in E$, we require approximately 
    $|V| + |E| + \sum_{e\in E} \log_2(\ell_e)$ bits. Note that we can 
    specify an edge with two pointers to the endpoints.
\end{itemize} 
We will need big-$O$ notation to describe running time, because we are 
interested in the asymptotic behaviour of algorithms. For two functions
$f : \R^+ \to \R^+$ and $g : \R^+ \to \R^+$, we say that $f(n) = 
O(g(n))$ if there exist constants $n_0 \in \N$ and $c \geq 0$ such that 
$f(n) \leq c \cdot g(n)$ for all $n \geq n_0$. For example, we have 
$2n^2 + 1 = O(n^2)$ and $\log(n) = O(n)$. An algorithm is then considered 
{\bf efficient} if its running time is bounded above by a polynomial function 
of the input size. 

Let's consider the running time of Dijkstra's algorithm by looking 
at a naive implementation. For ease of notation, we will write 
$|V| = n$ and $|E| = m$. Note that in the case that $G = (V, E)$ is connected, 
we have $m \geq n-1$. 
\begin{itemize}
    \item Step 1 can be performed using $O(n)$ operations since we are only
    assigning values for each vertex. 
    \item In Step 2, there are at most $n$ iterations. 
    \begin{itemize}
        \item Step 2.1 can be performed using $O(m)$ operations because 
        each edge will participate in at most two comparisons throughout the 
        entire iteration. 
        \item Step 2.2 takes $O(n)$ operations in order to determine the 
        vertex of minimum upper bound. 
    \end{itemize}
\end{itemize}
Therefore, the running time of the naive implementation is 
$O(n + mn + n^2) = O(mn)$, which is polynomial in the input size.  

We note that there are better implementations of Dijkstra's algorithm 
than the naive one that we have just stated. For Step 2.1, we can 
use $m$ \textsc{Decrease-Key} calls and for Step 2.2, we can use 
$n$ \textsc{Extract-Min} calls. In particular, by using Fibonacci heaps, 
\textsc{Decrease-Key} has running time $O(1)$ and \textsc{Extract-Min} 
has running time $O(\log n)$, which brings the total running time down to 
$O(m + n\log n)$.